Fail Faster

### 2025-03-24 paper/paper.tex
staging
wibble
CP

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  Property-based testing relies on random input generators to uncover bugs, often under tight time constraints. These generators are built 
  using generator libraries, which provide expressive combinators for constructing complex inputs. 
  Although convenient,
  today's generator libraries suffer from substantial performance costs, limited by abstraction overhead and inefficient randomness libraries.
  In this paper, we identify and address these bottlenecks. We introduce \name, a multi-stage programming technique that eliminates generator overhead
  without sacrificing the flexibility and expressiveness of existing generator library functions. We augment leading generator libraries in OCaml and 
  Scala 3 to use the \name approach, preserving the original library's semantics and enabling precise, pointwise performance comparisons. We also show that \rand
  has a measurable impact on generator performance, quantify its contribution, and show that speedups of $5.5\times$ are possible. Together, these 
  improvements lead to bug-finding up to $13\times$ faster---showing that high-level generator abstractions need not come at the cost of speed. \tr{Needs a lot of work. No spice.}
\end{abstract}

@@

\section{How can I Make My Generator Library Faster?}
\label{section:faster-generators}

>>>>

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
Property-based testing (PBT) relies on generators for random test
cases, often constructed using embedded domain specific languages,
which provide expressive combinators for building and composing
generators.  The effectiveness of PBT depends critically on the speed
of these generators.
%
However, careful measurements show that the generator performance of
two widely used PBT libraries falls well short of what is possible,
due principally to (1) the abstraction overhead of their
combinator-heavy style and (2) suboptimal sources of randomness.
We characterize, quantify, and address these bottlenecks.

To eliminate abstraction overheads, we propose a technique based on
multi-stage programming, dubbed \name.  We apply this technique to
leading generator libraries in OCaml and Scala 3, significantly
improving performance.
%
To quantify the performance impact of the randomness source, we carry
out a controlled experiment, replacing the randomness in the OCaml PBT
library with an optimized version.
%
Both interventions exactly preserve the semantics of generators,
enabling precise, pointwise comparisons.
%
Together, these improvements make bug finding up to $13\times$ faster.
\end{abstract}

@@

\section{How Can I Make My Generator Library Faster?}
\label{section:faster-generators}
